---
title: "GTAPViz: Utilities"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    css: mystyle.css
    number_sections: true
    code_folding: show
    self_contained: false
vignette: >
  %\VignetteIndexEntry{GTAPViz: Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)
```


```{r Dev Period, include = FALSE, eval = TRUE}
rm(list=ls())

try(devtools::load_all(".."), silent = TRUE)  # go up one level from /vignettes/

input_path <- system.file("extdata/in", package = "GTAPViz")
sl4.plot.data <- readRDS(file.path(input_path, "sl4.plot.data.rds"))
har.plot.data <- readRDS(file.path(input_path, "har.plot.data.rds"))
macro.data    <- readRDS(file.path(input_path, "macro.data.rds"))
```

This vignette provides other useful functions in the `GTAPViz` package.   


# Adding Unit and Description Columns {#sec:add-info}

If you manually import data or need to adjust the `"Description"` and `"Unit"` columns, you can use the following command:

```{r Add Info, eval = FALSE}
# Create mapping file
mapping_df <- data.frame(
  Variable = c("qgdp", "EV", "ppriv"),
  Description = c("Real GDP Index", "Welfare Equivalents", "Consumer Price Index"),
  Unit = c("Percent", "million USD", "percent"),
  stringsAsFactors = FALSE
)

datasets <- add_mapping_info(mapping_df, 
                             external_map = mapping_df, 
                             mapping = "Yes")
```


# Convert Different Units

You can convert all units across all data frames within the data list by using the following code:

```{r Converting Unit, eval = FALSE}
your_data <- sl4.plot.data   # Change this to your data list name

your_data <- convert_units(change_unit_from = c("BAHT", "million USD"),
                           change_unit_to = c("USD", "billion USD"),
                           adjustment = c("*34.5", "/1000"))
```

<details>
  <summary class="toggle-summary">Explanation</summary>
  
  <div class="tip-box">

**What does this code do?**

- Converts `BAHT` to `USD` by multiplying the values by 34.5.  

- Converts `million USD` to `billion USD` by dividing the values by 1,000.  

**Parameters:**

- `change_unit_from`: The original unit name to be changed in the `"Unit"` column.  
  If your dataset does not yet include a `"Unit"` column, please refer to [adding unit and description columns](#sec:add-info).

- `change_unit_to`: The new unit name to be recorded in the `"Unit"` column.  

- `adjustment`: The mathematical operation used to transform the values.  
  Accepts operators such as `/`, `*`, `+`, or `-`.

  </div>
</details>

# Renaming Columns and Data List Names {#sec:rename-column}

To modify column names across all datasets and/or rename data list names automatically—such as changing `REG` to `Region` or `COMM` to `Commodity`—use the following command:  

```{r Rename Column, eval = FALSE}
# Creating Mapping File
rename_col <- data.frame(
  old = c("REG", "COMM", "ACTS"),
  new = c("Region", "Commodity", "Activity")
  )

har.plot.data <- HARplus::rename_dims(har.plot.data, 
                                      mapping_df = rename_col,
                                      rename_list_names = FALSE)
```

<details>
  <summary class="toggle-summary">💡 Tip</summary>
  
<div class="tip-box">
You can create this mapping dataframe in Excel and import it into R. The column names must be **"old"** and **"new"**.  
</div>

</details>


# Sorting Output {#sec:sorting-output}

This function allows you to sort data before generating plots or tables.  
For example, you can display `EXP2` before `EXP1`, or `SEAsia` before `Oceania`.

It supports:

- Sorting by multiple columns simultaneously for precise output control 

- Applying consistent sorting across all data frames within a data list

See `?sort_plot_data` for full details.

Key arguments:  

- `cols` specifies the column(s) to sort based on a predefined order (e.g., in `sl4.plot.data`, `Experiment` and `Region` will follow the defined order).  

- `sort_by_value_desc` controls whether sorting is in descending order (`TRUE/FALSE/NULL`).   

To apply sorting, use the following command:  

```{r Sorting Data Function, eval = FALSE}
# Using the function
sort_sl4_plot <- sort_plot_data(
  sl4.plot.data,
  sort_columns = list(
    Experiment = c("EXP2", "EXP1"), # Column Name = Sorting Order
    Region = c("SEAsia", "Oceania")
    ),
  sort_by_value_desc = NULL
)
```


# Renaming Values {#sec:rename-value}

To automatically modify values in any column before plotting—such as renaming country names in the `REG` column (e.g., changing `"USA"` to `"United States"` or `"CHN"` to `"China"`)—use the following command:  

```{r Rename Value, eval = FALSE}
# Creating Mapping File
rename.region <- data.frame(
  ColumnName = "REG",
  OldName = c("USA", "CHN"),
  NewName = c("United States", "China"),
  stringsAsFactors = FALSE
)

sl4.plot.data_rename <- rename_value(sl4.plot.data, mapping.file = rename.region)
```

<br>
Another useful example is renaming the components of welfare decomposition before plotting.

```{r, eval = FALSE}
# Rename Value if needed
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

welfare_decomp <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)
```


# (HARplus) Manual Data Processing {#sec:manualdata}

This section is for non-GTAP models or manual data extraction for plotting. Follow these steps:

1. Use `"HARplus::load_sl4x"` or `"HARplus::load_harx"`, then `"HARplus::group_data_by_dims"` to structure the data in the required format.  

2. Filter each dataframe using standard R `data.frame` operations.  

3. Add the "Unit" and "Description" columns using `"add_mapping_info"` (the *"Unit"* column is required for plotting).  

```{r Manual Data Filtering, eval = FALSE}
# Step 1: Filtering Data using Dataframe and lapply
manual.data <- lapply(sl4data, function(x) {
  if (is.data.frame(x)) {
    x[x$REG %in% selected_regions & x$COMM %in% selected_sector & x$ACTS %in% selected_sector
      & x$Experiment %in% selected_exp , ]
  } else {
    x  
  }
})

# Step 2: Adding Unit and Description Column
manual.data <- add_mapping_info(manual.data, external_map = "/your/mapping.xlsx",
                                description_info = TRUE,
                                unit_info = TRUE)
```


# Sample Data

Sample data used in this vignette is obtained from the GTAPv7 model
and utilizes data from the GTAP 11 database. For
more details, refer to the [GTAP Database Archive](https://www.gtap.agecon.purdue.edu/databases/archives.asp).

```{=html}
<script>
window.addEventListener("load", function () {
  document.querySelectorAll("pre code").forEach(function (codeBlock) {
    const button = document.createElement("button");
    button.className = "copy-code-button";
    button.type = "button";
    button.innerText = "Copy";
    button.style.position = "absolute";
    button.style.top = "4px";
    button.style.right = "4px";

    button.addEventListener("click", () => {
      navigator.clipboard.writeText(codeBlock.innerText).then(() => {
        button.innerText = "Copied!";
        setTimeout(() => (button.innerText = "Copy"), 1500);
      });
    });

    const pre = codeBlock.parentNode;
    pre.style.position = "relative";
    pre.insertBefore(button, codeBlock);
  });
});
</script>

<script>
document.addEventListener("DOMContentLoaded", function () {
  let prevScrollPos = window.pageYOffset;
  const navbar = document.querySelector(".navbar");

  window.addEventListener("scroll", function () {
    const currentScrollPos = window.pageYOffset;
    if (prevScrollPos > currentScrollPos) {
      // Scrolling up
      navbar.style.top = "0";
    } else {
      // Scrolling down
      navbar.style.top = "-70px";
    }
    prevScrollPos = currentScrollPos;
  });
});
</script>
```
