---
title: 'GTAPViz: Plot Configurations'
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    css: mystyle.css
    number_sections: true
    code_folding: show
    self_contained: true
vignette: >
  %\VignetteIndexEntry{GTAPViz: Plot Configurations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)

```


```{r Dev Period, include = FALSE, eval = FALSE}
rm(list=ls())
#------------------Dev Period######################################
# Directory
project.folder <- "C:/Users/b_pat/OneDrive - purdue.edu/GTAPViz Data"
#project.folder <- "D:/One Drive/OneDrive - purdue.edu/GTAPViz Data"

input.folder <- paste0(project.folder, "/in")
map.folder <- paste0(project.folder, "/map")
output.folder <- paste0(project.folder, "/out")

sl4map <- readxl::read_xlsx(paste0(map.folder, "/OutputMapping.xlsx"), sheet = "SL4File")
harmap <- readxl::read_xlsx(paste0(map.folder, "/OutputMapping.xlsx"), sheet = "HARFile")

# 2. Experiment / Region / Sector Filter
experiment <- c("US_All", "US_All_RetalTar", "US_All_ReduceTar50", "US_All_RegReduceTar50",
                "US_All10", "US_All10_RetalTar", "US_All10_ReduceTar50", "US_All10_RegReduceTar50")


selected_regions <- c("USA", "CHN", "CAN", "ROW")
selected_sector <- NULL

# 3. Info Mode
info.mode <- "Mix"

# 4. Choosing Output: (CSV, STATA, R, TEXT) 
csv.output <- "no"    
stata.output <- "no"  
r.output <- "no"
txt.output <- "no"   

# 5. For Plotting: (TRUE/FALSE)
plot_data = TRUE

setwd <- "D:/GitHub/GTAPViz/R"
devtools::load_all()

# Extracting Data
auto_gtap_data(
  process_sl4_vars = sl4map, 
  process_har_vars = harmap,
  sl4_mapping_info = sl4map,
  har_mapping_info = harmap,
  region_select = selected_regions,
  sector_select = selected_sector,
  subtotal_level = FALSE,
  experiment = experiment, 
  mapping_info = info.mode,
  project_path = project.folder,
  plot_data = plot_data,
  output_formats = list(
    "csv" = csv.output,
    "stata" = stata.output,
    "rds" = r.output,
    "txt" = txt.output))

sl4.plot.data <- convert_units(sl4.plot.data, 
                               scale_auto = "mil2bil")

har.plot.data <- convert_units(har.plot.data, 
                               scale_auto = "mil2bil")

```

# Introduction {#sec:introduction}

This vignette is specifically designed to demonstrate how to simply apply all plot configurations within the package. For data extraction and full detail, please refer to [GTAPViz: Visualizing GTAP Results from HAR and SL4 Files](https://bodysbobb.github.io/GTAPViz/articles/GTAPViz_vignette.html).

This vignette will illustrate how to use:
- [Plot Style Configuraion](#sec:plot-config)

- [Color Palette](#sec:color-config)

- [Export Configuraion](#sec:export-config)

Before proceeding, ensure that `GTAPViz` is installed and loaded:
```{r}
library(GTAPViz)
```

Let's start by clarifying what these three configurations are by using a help function to call any plot function (any plot is fine since all of them have the same configuration parameters):

```{r}
?comparison_plot
```

```{r, eval=FALSE}
comparison_plot(
  data,
  filter_var = NULL,
  x_axis_from,
  split_by = NULL,
  panel_var = "Experiment",
  variable_col = "Variable",
  unit_col = "Unit",
  desc_col = "Description",
  invert_pane = FALSE,
  separate_figure = FALSE,
  var_name_by_description = FALSE,
  add_var_info = FALSE,
  output_path = NULL,
  export_picture = TRUE,
  export_as_pdf = FALSE,
  export_config = NULL,
  plot_style_config = NULL
)
```

# Plot Style Configuraion {#sec:plot-config}

To simply get all **plot configurations** using the following code:




<p> Firstly, multiple `color_tone` options are available in this package.  
You can view all predefined palettes by using the following code: </p>
```{r color palette, eval=FALSE}
# Get all palettes as a list
all_palettes <- get_color_palette("all")

# Click or call a specific palette to view its colors
all_palettes$winter("qualitative")   # View the winter palette
all_palettes$fall("sequential")                 # View the fall palette
all_palettes$gtap("diverging")     # View the GTAP palette
```

<p> **Note:** You can also define any other color palette simply by using `"blue"`, `"red"`, or any other standard color name that works with `ggplot2`. </p>

```{r, eval = FALSE}
get_all_config(printing = TRUE)
```

<p> You will get: </p>
```{r my style config, eval = FALSE}
my_style_config <- list(

  # Title settings
  show_title = TRUE,
  title_face = "bold",
  title_size = 20,
  title_hjust = 0.5,
  add_unit_to_title = TRUE,
  title_margin = margin(t = 10, r = 0, b = 10, l = 0),
  title_format = list(
    type = "standard", #option: prefix, suffix, full, dynamic
    text = "",
    sep = ""
  ),

  # X-Axis settings
  show_x_axis_title = TRUE,
  x_axis_title_face = "bold",
  x_axis_title_size = 16,
  x_axis_title_margin = margin(t = 25, r = 25, b = 0, l = 0),
  show_x_axis_labels = TRUE,
  x_axis_text_face = "plain",
  x_axis_text_size = 14,
  x_axis_text_angle = 0,
  x_axis_text_hjust = 0,
  x_axis_description = "",

  # Y-Axis settings
  show_y_axis_title = TRUE,
  y_axis_title_face = "bold",
  y_axis_title_size = 16,
  y_axis_title_margin = margin(t = 25, r = 25, b = 0, l = 0),
  show_y_axis_labels = TRUE,
  y_axis_text_face = "plain",
  y_axis_text_size = 14,
  y_axis_text_angle = 0,
  y_axis_text_hjust = 0,
  y_axis_description = "",
  show_axis_titles_on_all_facets = TRUE,

  # Value Labels
  show_value_labels = TRUE,
  value_label_face = "plain",
  value_label_size = 5,
  value_label_position = "above",
  value_label_decimal_places = 2,

  # Legend
  show_legend = FALSE,
  show_legend_title = FALSE,
  legend_position = "bottom",
  legend_title_face = "bold",
  legend_text_face = "plain",
  legend_text_size = 14,

  # Panel Strip
  strip_face = "bold",
  strip_text_size = 16,
  strip_background = "lightgrey",
  strip_text_margin = margin(t = 10, r = 0, b = 10, l = 0),

  # Panel Layout
  panel_spacing = 2,
  panel_rows = NULL,
  panel_cols = NULL,
  theme = NULL,

  # Colors
  color_tone = NULL,
  positive_color = "#2E8B57",
  negative_color = "#CD5C5C",
  background_color = "white",
  grid_color = "grey90",
  show_grid_major_x = FALSE,
  show_grid_major_y = FALSE,
  show_grid_minor_x = FALSE,
  show_grid_minor_y = FALSE,

  # Zero Line
  show_zero_line = TRUE,
  zero_line_type = "dashed",
  zero_line_color = "black",
  zero_line_size = 0.5,
  zero_line_position = 0,

  # Bar Chart
  bar_width = 0.9,
  bar_spacing = 0.9,

  # Scale Settings
  scale_limit = NULL,
  scale_increment = NULL,

  # Scale Expansion
  expansion_y_mult = c(0.05, 0.1),
  expansion_x_mult = c(0.05, 0.05),

  # Font Size Control
  all_font_size = 1,

  # Data Sorting
  sort_data_by_value = FALSE
)

# Example usage:
# comparison_plot(data, x_axis_from = "REG", 
#                 plot_style_config = my_style_config)
```

