---
title: "GTAPViz: Advanced Configurations and Visual Output Catalog"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    css: mystyle.css
    number_sections: true
    code_folding: show
    self_contained: true
runtime: shiny
---

```{r, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)

```


```{r Dev Period, include = FALSE}
rm(list=ls())
library(shiny)
library(DT)

setwd <- "D:/GitHub/GTAPViz/R"
devtools::load_all()

# 1. Project Directory
project.folder <- "D:/One Drive/OneDrive - purdue.edu/JGEA Paper/GTAPViz"

# 2. Define the Input Names 
experiment <- c("ReduceTar10","ReduceTar50", "IncreaseTar10","IncreaseTar50")

# 3. Adding Description / Unit (Yes/No/GTAPv7/Mix)
info.mode <- "Mix"

# 4. Choosing Output: (CSV, STATA, R, TEXT) 
csv.output <- "No"    
stata.output <- "No"  
r.output <- "No"
txt.output <- "No"   

# 5. For Plotting: (TRUE/FALSE)
plot_data = TRUE

# 1. Default Directory
input.folder <- paste0(project.folder, "/in")
map.folder <- paste0(project.folder, "/map")
output.folder <- paste0(project.folder, "/out")

# 2. Default Input Files
sl4map <- readxl::read_xlsx(paste0(map.folder, "/OutputMapping.xlsx"), sheet = "SL4File")
harmap <- readxl::read_xlsx(paste0(map.folder, "/OutputMapping.xlsx"), sheet = "HARFile")
filter.map <- readxl::read_xlsx(paste0(map.folder, "/OutputMapping.xlsx"), sheet = "FilterData")

# 3. Filter Data
selected_regions <- if(length(filter.map$Region) > 0) filter.map$Region else NULL
selected_sector  <- if(length(filter.map$Sector) > 0) filter.map$Sector else NULL

# 4. Extracting Data
auto_gtap_data(
  process_sl4_vars = sl4map, 
  process_har_vars = harmap,
  sl4_mapping_info = sl4map,
  har_mapping_info = harmap,
  region_select = selected_regions,
  sector_select = selected_sector,
  subtotal_level = FALSE,
  experiment = experiment, 
  mapping_info = info.mode,
  project_path = project.folder,
  add_scenario_ranking = TRUE,
  plot_data = plot_data,
  output_formats = list(
    "csv" = csv.output,
    "stata" = stata.output,
    "rds" = r.output,
    "txt" = txt.output))

# 5. Converting Units
sl4.plot.data <- convert_units(sl4.plot.data, 
                               scale_auto = "mil2bil")

har.plot.data <- convert_units(har.plot.data, 
                               scale_auto = "mil2bil")

region_data = sl4.plot.data[["1D"]][["Region"]]
sector_data = sl4.plot.data[["2D"]][["Sector"]]
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

har.plot.data <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)

# Extract Data
decomposition_data = har.plot.data[["A"]]

```

# Introduction {#sec:introduction}

This vignette is intended solely to illustrate all possible outputs for each function. For a comprehensive, step-by-step guide, please refer to the  
[GTAPViz – Vignette](https://bodysbobb.github.io/GTAPViz/articles/GTAPViz_plot.html).

# Data Sample

This vignette illustrates all output formats from each function using results from the standard GTAPv7 model and the GTAP 11c database. The dataset used in this example has the following characteristics:

- 10×10 aggregation (i.e., 10 regions and 10 sectors)  
- Four distinct simulations (experiments): `"ReduceTar10"`, `"ReduceTar50"`, `"IncreaseTar10"`, and `"IncreaseTar50"`  
- Variable units and descriptions follow GTAPv7 defaults

The dataset is generated using the `auto_gtap_data()` function. For step-by-step instructions on constructing this dataset, please refer to the [GTAPViz – Vignette](https://bodysbobb.github.io/GTAPViz/articles/GTAPViz_plot.html).

**Note:** The **GTAP model and database** require a separate license, which can be obtained at the [GTAP DataBase](https://www.gtap.agecon.purdue.edu/databases/pricing.asp) website.


## Region Data

`region_data` is one-dimensional dataset (i.e., `REG`) is used for the `comparison_plot` function. It contains:

- `REG`: 10 regions  
- `Variable`: `qgdp`, `ppriv`, `u`, `EV`

<details>
<summary class="toggle-summary">Click to show/hide `region_data` </summary>

```{r region_data, eval=FALSE}
region_data = sl4.plot.data[["1D"]][["Region"]]
```


```{r region_table, echo=FALSE}
# UI: Dropdown filter
selectInput("var_filter", 
            label = HTML("<span class='select-label'>Choose a Variable to Filter:</span>"),
            choices = unique(region_data$Variable),
            selected = unique(region_data$Variable)[1],
            width = '300px')

# Server: Filtered DataTable
DT::renderDataTable({
  filtered_data <- subset(region_data, Variable == input$var_filter)

  datatable(
    filtered_data,
    filter = "top",
    options = list(
      pageLength = 20,
      lengthMenu = c(10, 20, 50),
      autoWidth = TRUE,
      scrollX = TRUE,
      dom = 'tip',
      columnDefs = list(
        list(className = 'dt-center', targets = "_all")
      )
    ),
    rownames = FALSE,
    class = 'compact cell-border stripe'
  )
})
```
</details>

## Sector Data

`sector_data` is two-dimensional dataset (i.e., `REG*COMM`) is used for the `detail_plot` function. It contains:

- `REG`: 10 regions  
- `COMM`: 10 sectors  
- `Variable`: `qo`, `qxw`

<details>
<summary class="toggle-summary">Click to show/hide `sector_data` </summary>

```{r sector_data, eval=FALSE}
sector_data = sl4.plot.data[["2D"]][["Sector"]]
```

```{r sector_table, echo=FALSE}
# UI: Dropdown filter for sector-level data
selectInput("sector_var_filter", 
            label = HTML("<span class='select-label'>Choose a Variable to Filter:</span>"),
            choices = unique(sector_data$Variable),
            selected = unique(sector_data$Variable)[1],
            width = '300px')

# Server: Filtered DataTable
DT::renderDataTable({
  filtered_data <- subset(sector_data, Variable == input$sector_var_filter)

  datatable(
    filtered_data,
    filter = "top",
    options = list(
      pageLength = 20,
      lengthMenu = c(10, 20, 50),
      autoWidth = TRUE,
      scrollX = TRUE,
      dom = 'tip',
      columnDefs = list(
        list(className = 'dt-center', targets = "_all")
      )
    ),
    rownames = FALSE,
    class = 'compact cell-border stripe'
  )
})
```
</details>

## Decomposition Data

`decomposition_data` contains a welfare decomposition derived from GTAP model results, specifically from the `welfare.har` file under header `A`. It is used for the `stack_plot` function and includes:

- `REG`: 10 regions  
- `COLUMN`: Components of welfare (for decomposition)

**Note:** This dataset may optionally require renaming of component values to enhance figure readability and clarity.

<details>
<summary class="toggle-summary">Click to show/hide `decomposition_data` </summary>

```{r decomposition_data, eval=FALSE}
# Rename Components
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

har.plot.data <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)

# Extract Data
decomposition_data = har.plot.data[["A"]]
```


```{r decomposition_table, echo=FALSE}
DT::renderDataTable({
  datatable(
    decomposition_data,
    filter = "top",  # still enables filters for all columns
    options = list(
      pageLength = 20,
      lengthMenu = c(10, 20, 50),
      autoWidth = TRUE,
      scrollX = TRUE,
      dom = 'tip',
      columnDefs = list(
        list(className = 'dt-center', targets = "_all")
      )
    ),
    rownames = FALSE,
    class = 'compact cell-border stripe'
  )
})
```
</details>
<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>

# Advanced Configurations {#sec:plot-title-and-output-name}

## Plot Styles

This package allows users to fully customize plot appearance based on the `ggplot2` package. Style preferences can be defined once and stored as a list object in the R environment, enabling consistent application across all plotting functions.

To preview the default style configuration, run the following:

<details>
<summary class="toggle-summary">Click to show/hide the `plot_style_config` </summary>

```{r Plot Style Config, echo=TRUE}
get_plot_style_config(printing = TRUE)
```

</details>

## Plot Titles

### Default Plot Title

By default, plot titles are **auto-generated** using values from the following columns. These can be redefined as needed to match the structure of your dataset:

- `variable_col`: the variable name (default: `"Variable"`)

- `split_by`: the column used to split plots (default: `NULL`)

- `unit_col`: the unit of measurement (default: `"Unit"`)

- `panel_var`: the experiment name (used when `separate_figure = TRUE`, default: `"Experiment"`)

<details>
<summary class="toggle-summary">💡 Tip: Plot Title</summary>

<div class="tip-box">
1. `var_name_by_description` replaces the variable name (`variable_col`) with its description from `desc_col`. <br>
2. `add_var_info` controls whether the plot title displays the variable name, its description, or both (e.g., "Real GDP (qgdp)").
</div>
</details>

### Plot Title Format

`title_format`, defined within `plot_style_config`, is specifically designed to modify the plot title.


```{r Plot Title Format, echo=TRUE, eval=FALSE}
plot_style_config = list(
  
  title_format = list(
    type = "standard", # option: prefix, suffix, full, dynamic
    text = "",
    sep = ""
  )
)
```

Users can define `title_format` using one of five title types:

- `"standard"` *(default)*: Generates the title automatically based on the variable, description, and unit.  

- `"prefix"`: Adds a user-defined prefix (`text`) to the auto-generated title, separated by `sep`.  

- `"suffix"`: Adds a user-defined suffix (`text`) to the auto-generated title, separated by `sep`.  

- `"full"`: Replaces the auto-generated title entirely with the specified `text`.  

- `"dynamic"`: Creates a custom title from a template combining column values and user-defined text.

<details>
<summary class="toggle-summary">💡 Tip: Plot Title Format</summary>

<div class="tip-box">
It is highly recommended to include `add_unit_to_title` in `plot_style_config`.

```{r Plot Title Format with Unit, echo=TRUE, eval=FALSE}
plot_style_config = list(
  title_format = list(
    type = "standard",
    text = "",
    sep = ""
  ),
  add_unit_to_title = TRUE
)
```
</div>

</details>

```{r dynamic title, include=FALSE, echo=FALSE}
dynamic_title <- region_data
dynamic_title <- subset(dynamic_title, select = -c(Dimension, ScenarioRank, Value))
dynamic_title <- dynamic_title[order(dynamic_title$Region), ]
dynamic_title[["Default Title"]] <- glue::glue_data(dynamic_title, "{Variable} ({Unit})")
dynamic_title[["Prefix Title"]] <- glue::glue_data(dynamic_title, "Impact on {Variable} ({Unit})")
dynamic_title[["Suffix Title"]] <- glue::glue_data(dynamic_title, "{Variable} ({Unit}) Impacts")
dynamic_title[["Full Title"]] <- glue::glue_data(dynamic_title, "Impact on Variables in Units")
dynamic_title[["Dynamic Title"]] <- glue::glue_data(dynamic_title, "Impact on {Variable} in {Region} ({Unit})")
dynamic_title[["Auto Add Unit to Title"]] <- glue::glue_data(
  dynamic_title,
  "Impact on {Variable} in {Region} ({ifelse(tolower(Unit) == 'percent', '%', Unit)})"
)
dynamic_title <- dynamic_title[1:40, ]
```

The table below shows example plot titles; the last column contains the titles that would be used in the visualizations, assuming the plot function is called as follows:

```{r Plot sample for title, eval=FALSE}
comparison_plot(data = region_data,
                x_axis_from = "Region",  
                split_by = "Variable",
                panel_var = "Experiment",
                variable_col = "Variable",
                unit_col = "Unit",
                desc_col = "Description",
```


```{r plot-title-ui, echo=FALSE}
# UI Dropdown
selectInput("title_style", "Select Plot Title Format:",
            choices = c("Default", "Prefix/Suffix", "Full", "Dynamic", "Auto Add Unit to Title"),
            selected = "Default", width = "300px")
```

```{r plot-title-outputs, echo=FALSE}
# Reactive outputs must be defined inside a Shiny chunk
observe({
  output$title_explainer <- renderUI({
    if (input$title_style == "Dynamic") {
      tags$p("Creates a customized title using a template string, where you can insert values from your data frame columns using curly braces `{}`.")
    } else if (input$title_style == "Auto Add Unit to Title") {
      tags$p("I highly recommend using the `add_unit_to_title` option under `plot_style_config` instead of manually adding the `Unit` column to the title. This option automatically converts \"Percent\" to the `%` symbol and wraps it in parentheses.")
    }
  })

  output$title_code <- renderUI({
    code_block <- switch(input$title_style,
      "Default" = 'title_format = list(\n  type = "standard"\n)',
      "Prefix/Suffix" = '# Prefix Sample\ntitle_format = list(\n  type = "prefix",\n  text = "Impact on",\n  sep = ""\n)\n\n# Suffix Sample\ntitle_format = list(\n  type = "suffix",\n  text = "Impacts",\n  sep = ""\n)',
      "Full" = 'title_format = list(\n  type = "full",\n  text = "Impact on Variables in Units",\n  sep = ""\n)',
      "Dynamic" = 'title_format = list(\n  type = "full",\n  text = "Impact on {Variable} in {Region} ({Unit})",\n  sep = ""\n)',
      "Auto Add Unit to Title" = 'plot_style_config = list(\n  title_format = list(\n    type = "dynamic",\n    text = "Impact on {Variable} in {Region}",\n    sep = ""\n  ),\n  add_unit_to_title = TRUE\n)'
    )

    tags$pre(style = "background-color:#f8f9fa; border:1px solid #ccc; padding:10px; font-family: monospace; white-space: pre-wrap;",
             code_block)
  })

  output$title_notes <- renderUI({
    if (input$title_style == "Dynamic") {
      tags$ul(
        tags$li("{Variable}, {Region}, and {Unit} are placeholders that will be replaced with the corresponding values from each observation."),
        tags$li('Fixed text such as "Impact on" and "in" remains constant and will appear in all plot titles.')
      )
    }
  })

  output$title_table <- DT::renderDataTable({
    show_cols <- switch(input$title_style,
      "Default" = c("Region", "Experiment", "Variable", "Unit", "Description", "Default Title"),
      "Prefix/Suffix" = c("Region", "Experiment", "Variable", "Unit", "Description", "Prefix Title", "Suffix Title"),
      "Full" = c("Region", "Experiment", "Variable", "Unit", "Description", "Full Title"),
      "Dynamic" = c("Region", "Experiment", "Variable", "Unit", "Description", "Dynamic Title"),
      "Auto Add Unit to Title" = c("Region", "Experiment", "Variable", "Unit", "Description", "Auto Add Unit to Title")
    )

    DT::datatable(
      dynamic_title[, show_cols],
      filter = "none",
      options = list(
        pageLength = 10,
        autoWidth = TRUE,
        scrollX = TRUE,
        dom = 'tip',
        columnDefs = list(list(className = 'dt-left', targets = "_all"))
      ),
      rownames = FALSE,
      class = 'compact cell-border stripe small-font'
    )
  })
})
```

```{r plot-title-uioutputs, echo=FALSE}
uiOutput("title_explainer")
uiOutput("title_code")
uiOutput("title_notes")
DT::dataTableOutput("title_table")
```



## Exported Figure

Exported figures are automatically named to reflect their corresponding plot titles, ensuring clarity and differentiation across outputs. Additional export settings—such as background color and image quality—can be customized using the `export_config` list. For full documentation, run:

<details>
<summary class="toggle-summary">Click to show/hide the `export_config` </summary>

```{r Export Config, echo=TRUE}
get_export_config(printing = TRUE)
```

</details>

## Implementation

Once you have defined and saved your plot style settings, you can store multiple configuration lists and apply them to any plotting function by passing the saved list as an argument.

<details>
<summary class="toggle-summary">Click to show/hide the sample code for applying configurations</summary>

```{r Implementation Config, echo=TRUE, eval=FALSE}
comparison_plot(data = region_data,
                x_axis_from = "Region",
                split_by = "Variable",

                # Assumes that my_export_config contains export settings
                export_config = my_export_config,

                # Assumes that my_style_config contains plot style settings
                plot_style_config = my_style_config)
```

</details>


<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>

<div class="info-box">
<strong>Important:</strong> <br>
  Output samples shown in this vignette may appear slightly distorted due to HTML rendering. This is not an issue in actual usage, where output dimensions are controlled by the `width`, `height`, and `dpi` options in `export_config`.
</div>

<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>


# Comparison Plot {#sec:comparison-plot}

All examples of the `comparison_plot` function in this section use one-dimensional data (`region_data`) for visualization.

## Example 1 (Default)

This example demonstrates the default function settings:

- `x_axis_from = "Region"`: The x-axis displays all 10 regions.  
  *Note:* If the `auto_gtap_data()` function is not used, this column is labeled `REG` in the standard GTAP output.  
  
- `split_by = "Variable"`: A separate panel is created for each variable. Since the dataset contains four variables, four subplots will be generated.  

- `panel_var = "Experiment"`: Specifies the simulation scenarios, consisting of four experiments.  

- `invert_pane = FALSE`: The x-axis is displayed horizontally.  

- `separate_figure = FALSE`: All variables are displayed within a single figure using a multi-panel layout, rather than generating separate figures.

```{r Comparison Plot Sample}
plotA <- comparison_plot(data = region_data,
                          filter_var = NULL,    
                          x_axis_from = "Region",  
                          split_by = "Variable",
                          panel_var = "Experiment",
                          variable_col = "Variable",
                          unit_col = "Unit",
                          desc_col = "Description",
                          invert_pane = FALSE,
                          separate_figure = FALSE,
                          var_name_by_description = FALSE,
                          add_var_info = FALSE,
                          output_path = output.folder,
                          export_picture = FALSE,
                          export_as_pdf = FALSE,
                        
                          # See ?export_config
                          export_config = list(
                            width = 20,
                            height = 12
                            ),
                          
                          # See ?plot_style_config
                          plot_style_config = list(
                            color_tone = "gtap",
                            add_unit_to_title = TRUE,
                            title_format = list(
                              type = "prefix",  
                              text = "Impact on"
                            ),
                            panel_rows = 2,
                            bar_width = 0.4,
                            bar_spacing = 0.6,
                            x_axis_text_angle = 45,
                            plot.margin = c(10, 50, 10, 10)
                          )
                        )
```

For this example, the total number of output plots is calculated as `split_by = 4`, resulting in **4 figures** being generated.

```{r interactive-plotAEXP1, echo=FALSE}
selectInput("plot_choiceA1", "Example 1 - Outputs", choices = names(plotA))
uiOutput("plot_title")

renderUI({
  HTML(paste0("<b>Exported Filename:</b> ", input$plot_choiceA1))
})

renderPlot({
  plotA[[input$plot_choiceA1]]
}, width = 1600, height = 860)
```


## Example 2 (`invert_pane`)

Same as example 1 only that `invert_pane = TRUE`

**The number of plots generated in this example matches that of Example 1.**

<details>
<summary class="toggle-code-summary">[Click to show/hide code]</summary>
```{r PlotA2}
plotA2 <- comparison_plot(data = region_data,
                          filter_var = NULL,    
                          x_axis_from = "Region",  
                          split_by = "Variable",
                          panel_var = "Experiment",
                          variable_col = "Variable",
                          unit_col = "Unit",
                          desc_col = "Description",
                          invert_pane = TRUE,
                          separate_figure = FALSE,
                          var_name_by_description = FALSE,
                          add_var_info = FALSE,
                          output_path = NULL,
                          export_picture = FALSE,
                          export_as_pdf = FALSE,
                        
                          # See ?export_config
                          export_config = list(
                            width = 20,
                            height = 12
                            ),
                          
                          # See ?plot_style_config
                          plot_style_config = list(
                            color_tone = "gtap",
                            add_unit_to_title = TRUE,
                            title_format = list(
                              type = "prefix",  
                              text = "Impact on"
                            ),
                            panel_rows = 2,
                            bar_width = 0.4,
                            bar_spacing = 0.6,
                            x_axis_text_angle = 45,
                            plot.margin = c(10, 50, 10, 10)
                          )
                        )
```

</details>

```{r interactive-plotAEXP2, echo=FALSE}
selectInput("plot_choiceA2", "Example 2 - Outputs", choices = names(plotA2))
uiOutput("plot_title")

renderUI({
  HTML(paste0("<b>Exported Filename:</b> ", input$plot_choiceA2))
})

renderPlot({
  plotA2[[input$plot_choiceA2]]
}, width = 1600, height = 860)
```


## Example 3 (`separate_figure`)

This example is identical to Example 1, except that `separate_figure = TRUE`.  

In this case, the output will be automatically split by `panel_var` (which defaults to the `"Experiment"` column), calculated as `split_by` × `panel_var` = `4 × 4`, resulting in **16 figures** being generated.

<details>
<summary class="toggle-code-summary">[Click to show/hide code]</summary>
```{r plotA3}
plotA3 <- comparison_plot(data = region_data,
                          filter_var = NULL,    
                          x_axis_from = "Region",  
                          split_by = "Variable",
                          panel_var = "Experiment",
                          variable_col = "Variable",
                          unit_col = "Unit",
                          desc_col = "Description",
                          invert_pane = FALSE,
                          separate_figure = TRUE,
                          var_name_by_description = FALSE,
                          add_var_info = FALSE,
                          output_path = NULL,
                          export_picture = FALSE,
                          export_as_pdf = FALSE,
                        
                          # See ?export_config
                          export_config = list(
                            width = 20,
                            height = 12
                            ),
                          
                          # See ?plot_style_config
                          plot_style_config = list(
                            color_tone = "gtap",
                            add_unit_to_title = TRUE,
                            title_format = list(
                              type = "prefix",  
                              text = "Impact on"
                            ),
                            panel_rows = 2,
                            bar_width = 0.4,
                            bar_spacing = 0.6,
                            x_axis_text_angle = 45,
                            plot.margin = c(10, 50, 10, 10)
                          )
                        )
```
</details>

```{r interactive-plotAEXP3, echo=FALSE}
selectInput("plot_choiceA3", "Example 3 - Outputs", choices = names(plotA3), width = '500px')
uiOutput("plot_title")

renderUI({
  HTML(paste0("<b>Exported Filename:</b> ", input$plot_choiceA3))
})

renderPlot({
  plotA3[[input$plot_choiceA3]]
}, width = 1600, height = 860)
```

<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>


# Detail Plot {#sec:detail-plot}

All examples of the `detail_plot` function in this section use two-dimensional data (`sector_data`) for visualization.

## Example 4 (`filter_var`)

This example demonstrates the benchmark settings for `detail_plot`:

- `filter_var = "qo"`: Plots only the `qo` variable, excluding `qxw`.  

- `x_axis_from = "Sector"`: The x-axis represents all 10 sectors.  

- `split_by = "Region"`: Generates a separate panel for each region. With 10 regions, the result is 10 subplots.  

- `panel_var = "Experiment"`: Defines the simulation scenarios, consisting of four experiments.  

- `invert_pane = TRUE`: Displays the x-axis vertically (i.e., treated as the y-axis), improving readability.  

- `separate_figure = FALSE`: All variables are displayed in a single figure using a multi-panel layout, rather than generating separate figures.

**Optional:**  

- `var_name_by_description = TRUE`: Use the description instead of the variable name in the plot title.  

- `add_var_info = TRUE`: Append the GTAP variable name in parentheses.

<details>
<summary class="toggle-summary">💡 Tip: Detail Plot</summary>

<div class="tip-box">
<strong>Tip:</strong><br>
1. Use `invert_pane = TRUE` to improve readability, especially for complex components like `COMM` and `ACTS`.<br>
2. The plot supports `positive_color` and `negative_color` to visually distinguish value signs; colors are automatically shaded based on value intensity.
</div>

</details>

```{r Detail Plot}
plotB <- detail_plot(data = sector_data,
                     filter_var = "qo",
                     x_axis_from = "Sector",
                     split_by = "Region",
                     panel_var = "Experiment",
                     variable_col = "Variable",
                     unit_col = "Unit",
                     desc_col = "Description",
                     var_name_by_description = TRUE,
                     add_var_info = TRUE,

                     top_impact = NULL,
                     separate_figure = FALSE,
                     invert_pane = TRUE,
                     
                     output_path = output.folder,
                     export_picture = FALSE,
                     export_as_pdf = FALSE,
                     export_config = list(
                       width = 25,
                       height = 20),
                     
                     plot_style_config = list(
                       positive_color = "#2E8B57", 
                       negative_color = "#CD5C5C",
                       panel_rows = 1,
                       panel_cols = NULL,
                       show_axis_titles_on_all_facets = FALSE,
                       bar_width = 0.6,
                       all_font_size = 1.2
                       )
                     )

```


For this example, the total number of output plots is calculated as `split_by × variable_col = 10 × 1`, resulting in **10 figures** being generated.

```{r interactive-plotBEXP4, echo=FALSE}
selectInput("plot_choiceB", "Example 4 - Outputs", choices = names(plotB), width = '500px')
uiOutput("plot_title")

renderUI({
  HTML(paste0("<b>Exported Filename:</b> ", input$plot_choiceB))
})

renderPlot({
  plotB[[input$plot_choiceB]]
}, width = 1600, height = 860)
```


## Example 5 (`top_impact`)

Same as example 4 only that `top_impact = 5`

**The number of plots generated in this example matches that of Example 4.**

<details>
<summary class="toggle-code-summary">[Click to show/hide code]</summary>

```{r Detail PlotB2}
plotB2 <- detail_plot(data = sector_data,
                     filter_var = "qo",
                     x_axis_from = "Sector",
                     split_by = "Region",
                     panel_var = "Experiment",
                     variable_col = "Variable",
                     unit_col = "Unit",
                     desc_col = "Description",
                     var_name_by_description = TRUE,
                     add_var_info = TRUE,

                     top_impact = 5,
                     separate_figure = FALSE,
                     invert_pane = TRUE,
                     
                     output_path = output.folder,
                     export_picture = FALSE,
                     export_as_pdf = FALSE,
                     export_config = list(
                       width = 25,
                       height = 20),
                     
                     plot_style_config = list(
                       positive_color = "#2E8B57", 
                       negative_color = "#CD5C5C",
                       panel_rows = 1,
                       panel_cols = NULL,
                       show_axis_titles_on_all_facets = TRUE,
                       bar_width = 0.6,
                       all_font_size = 1.2
                       )
                     )

```

</details>

```{r interactive-plotBEXP5, echo=FALSE}
selectInput("plot_choiceB2", "Example 5 - Outputs", choices = names(plotB2), width = '500px')
uiOutput("plot_title")

renderUI({
  HTML(paste0("<b>Exported Filename:</b> ", input$plot_choiceB2))
})

renderPlot({
  plotB2[[input$plot_choiceB2]]
}, width = 1600, height = 860)
```

<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>

# Stack Plot {#sec:stack-plot}

```{r Rename Decomposition File, eval = FALSE}
# Rename Value if needed
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

har.plot.data <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)

# Rename Column
rename_col <- data.frame(
  old = c("REG", "COMM", "ACTS"),
  new = c("Region", "Commodity", "Activity")
  )

har.plot.data <- HARplus::rename_dims(har.plot.data, rename_col)
```

## Stack plot

```{r Stack Plot, eval = FALSE}
stack_plot(data = har.plot.data[["A"]],
           x_axis_from = "Region",
           stack_value_from = "COLUMN",
           split_by = FALSE,
           
           show_total = TRUE,
           unstack_plot = FALSE,
           
           var_name_by_description = TRUE,       
           
           invert_pane = FALSE,           
           separate_figure = FALSE,

           export_picture = TRUE,
           export_as_pdf = "merged",
           export_config = list(
             width = 28,
             height = 18
           ),
           output_path = output.folder,
           
           plot_style_config = list(
             color_tone = "gtap",
             panel_rows = 2,
             panel_cols = NULL,
             show_legend = TRUE,
             show_axis_titles_on_all_facets = FALSE,
             all_font_size = 1.2,
             x_axis_text_angle = 45
           ))
```

## Unstack plot

```{r Unstack Plot, eval = FALSE}
stack_plot(data = har.plot.data[["A"]],
           x_axis_from = "Region",
           stack_value_from = "COLUMN",
           split_by = FALSE,
           
           show_total = TRUE,
           unstack_plot = TRUE,
           
           var_name_by_description = TRUE,       
           
           invert_pane = FALSE,           
           separate_figure = FALSE,

           export_picture = TRUE,
           export_as_pdf = FALSE,
           export_config = list(
             width = 28,
             height = 18
           ),
           output_path = output.folder,
           
           plot_style_config = list(
             color_tone = "gtap",
             panel_rows = 2,
             panel_cols = NULL,
             show_legend = TRUE,
             show_axis_titles_on_all_facets = FALSE,
             all_font_size = 1.2,
             x_axis_text_angle = 45
           ))
```


<div style="border-top: 1px solid #003366; margin-top: 15px; margin-bottom: 15px;"></div>

# Report Table {#sec:comparison-table}

## One-Dimensional Data  
A dataset structured by a single dimension, such as `REG`, `COMM`, or `ACTS`.
```{r 1D Table, eval = FALSE}
report_table(
  data_list = sl4.plot.data[["1D"]],
  pivot_col = list(Region = "Variable"),
  group_by =  list(
    Region = list("Experiment", "Region")),
  rename_cols = list("Experiment" = "Scenario"),

  total_column = FALSE,
  decimal = 4,
  subtotal_level = FALSE,
  repeat_label = FALSE,
  include_units = TRUE,

  var_name_by_description = TRUE,
  add_var_info = TRUE,
  add_group_line = FALSE,
  
  separate_sheet_by = "Unit",
  export_table = TRUE,
  output_path = output.folder,
  separate_file = FALSE,
  workbook_name = "Comparison Table 1D"
)
```

## Two-Dimensional Data  
A dataset structured by two dimensions, such as `REG*COMM` or `REG*ACTS`.
```{r 2D Table, eval = FALSE}
report_table(
  data_list = sl4.plot.data[["2D"]],
  pivot_col = list(Sector = "Variable"),
  group_by =  list(
    Sector = list("Experiment", "Region", "Sector")),
  rename_cols = list("Experiment" = "Scenario"),

  total_column = FALSE,
  decimal = 4,
  subtotal_level = FALSE,
  repeat_label = FALSE,
  include_units = TRUE,

  var_name_by_description = TRUE,
  add_var_info = TRUE,
  add_group_line = FALSE,
  
  separate_sheet_by = "Unit",
  export_table = TRUE,
  output_path = output.folder,
  separate_file = FALSE,
  workbook_name = "Comparison Table 2D"
)
```

## Multi-Dimensional Data  
A dataset involving more than two dimensions, such as `COMM*REG*REG`.
```{r 3D Table, eval = FALSE}
bilateral_data <- list(BiTrade = bilateral_data)
report_table(
  data_list = sl4.plot.data[["1D"]],
  pivot_col = list(Region = "Variable"),
  group_by =  list(
    Region = list("Experiment", "Region")),
  rename_cols = list("Experiment" = "Scenario"),

  total_column = FALSE,
  decimal = 4,
  subtotal_level = FALSE,
  repeat_label = FALSE,
  include_units = TRUE,

  var_name_by_description = TRUE,
  add_var_info = TRUE,
  add_group_line = FALSE,
  
  separate_sheet_by = "Unit",
  export_table = TRUE,
  output_path = output.folder,
  separate_file = FALSE,
  workbook_name = "Comparison Table"
)
```

# Pivot Table
```{r Pivot Table with Filter, eval=FALSE}
pivot_table_with_filter(
  data = sl4.plot.data[["2D"]][["Sector"]],
  filter = c("Variable", "Unit"),    # Allow filtering by variable type and unit
  rows = c("Region", "Sector"),      # Regions and sectors as row fields
  cols = c("Experiment"),           # Experiments as column fields
  data_fields = "Value",            # Values to be aggregated
  raw_sheet_name = "Raw_Data",     # Sheet name for raw data
  pivot_sheet_name = "Sector_Pivot", # Sheet name for pivot table
  dims = "A3",                      # Starting cell for pivot table
  export = TRUE,
  output_path = output.folder,
  workbook_name = "Sectoral_Impact_Analysis.xlsx"
)
```

