---
title: "Plot: A Step-by-Step Guide"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    css: mystyle.css
    number_sections: true
    code_folding: show
    self_contained: false
vignette: >
  %\VignetteIndexEntry{Plot: A Step-by-Step Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)
```


```{r Dev Period, include = FALSE, eval = TRUE}
try(devtools::load_all(".."), silent = TRUE)  # go up one level from /vignettes/

input_path <- system.file("extdata/in", package = "GTAPViz")
sl4.plot.data <- readRDS(file.path(input_path, "sl4.plot.data.rds"))
bilateral_data <- readRDS(file.path(input_path, "bilateral_data.rds"))
har.plot.data <- readRDS(file.path(input_path, "har.plot.data.rds"))
macro.data    <- readRDS(file.path(input_path, "macro.data.rds"))
```

This vignette illustrates how to set up plot configurations and generate outputs.  
For a complete list of available plot types with sample images, see the [plot catalog](https://pattawee.shinyapps.io/gtapviz-advanced-plot-configs/).

# Comparison Plot {#sec:comparisonplot}

This figure compares one or multiple variables across all experiments for selected observations from a chosen dimension (e.g., region, sector).  

<strong style="font-size: 1.17em;">Key Features</strong>

- Designed for comparing selected variables across experiments.  

- Displays all selected variables in one plot, making it ideal for a few observations; too many can make it dense and unclear.  

## Example 1

Plotting `qgdp`, `ppriv`, `EV`, `tot`, and `u` generates six separate figures using the following command:  

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
reg_data = sl4.plot.data[["REG"]]

plotA <- comparison_plot(
          # === Input Data ===
          data         = reg_data,
          filter_var   = NULL,
          x_axis_from  = "Region",
          split_by     = "Variable",
          panel_var    = "Experiment",
          variable_col = "Variable",
          unit_col     = "Unit",
          desc_col     = "Description",
        
          # === Plot Behavior ===
          invert_axis     = FALSE,
          separate_figure = FALSE,
        
          # === Variable Display ===
          var_name_by_description = FALSE,
          add_var_info            = FALSE,
        
          # === Export Settings ===
          output_path     = NULL,
          export_picture  = FALSE,
          export_as_pdf   = FALSE,
          export_config   = create_export_config(
            width  = 20,
            height = 12
          ),
        
          # === Styling ===
          plot_style_config = create_plot_style(
            color_tone         = "purdue",
            add_unit_to_title  = TRUE,
            title_format       = create_title_format(
              type = "prefix",
              text = "Impact on"
            ),
            panel_rows = 2
          )
        )
```

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(plotA[[1]])
```
</div> <figcaption class="package-fig"> Comparison Plot Visualization <br> </figcaption>
<span class="package-fig-note">(splt_by = FALSE, panel_rows = 2, panel_cols = NULL i.e., default)</span> </figcaption> </figure>


## GTAP Macro Variable Plot {#sec:macroplot}

This figure displays aggregated values across all experiments, typically for global economic impacts. It is designed for GTAP Macro variables extracted from the *"Macros"* header.

<strong style="font-size: 1.17em;">Key Features</strong>

- Compares macroeconomic indicators across experiments.  

- Works with aggregated data, making it ideal for global-scale analysis.  

- Primarily used for GTAP Macro variables but can be applied to other aggregated datasets.  

To plot this figure—the only difference from previous figures is changing `split_by = FALSE`—you can use the following command:

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
macro.plot <- comparison_plot(
  # === Input Data ===
  data         = macro.data[["macros"]],
  filter_var   = c("pgdpwld", "qgdpwld", "vgdpwld"),
  x_axis_from  = "Variable",
  split_by     = FALSE,     # THIS IS THE MOST IMPORTANT PART FOR THIS PLOT

  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 20,
    height = 15
  ),

  # === Styling ===
  plot_style_config = create_plot_style(
    color_tone = "gtap",
    title_format = create_title_format(
      type = "full",
      text = "Global Economic Impacts"
    )
  )
)
```

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(macro.plot[[1]])
```
</div> <figcaption class="package-fig"> GTAP Macro Plot Visualization <br> <span class="package-fig-note">(splt_by = FALSE, panel_rows and panel_cols = NULL i.e., default)</span> </figcaption> </figure>


# Detail Plot {#sec:detailplot}

This figure visualizes large-scale data, displaying all sectors (*COMM*, *ACTS*) or all regions (*REG*) within each experiment.  

<strong style="font-size: 1.17em;">Key Features</strong>

- Supports visualization across two selected dimensions.  

- Suitable for datasets too large for *comparison_plot*.  

- Automatically detects and aligns variables with different dimension names.  

- Filters data by selecting the most affected variables using **top_impact**.  

## Example 1

The following command plots all variables with sectors on the x-axis for each region:  

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
sector_data <- sl4.plot.data[["COMM*REG"]]

plotB <- detail_plot(
  # === Input Data ===
  data        = sector_data,
  filter_var  = list(Region = c("Oceania")),
  x_axis_from = "Commodity",
  split_by    = "Region",
  panel_var   = "Experiment",
  variable_col = "Variable",
  unit_col     = "Unit",
  desc_col     = "Description",
  
  # === Plot Behavior ===
  invert_axis      = TRUE,
  separate_figure  = FALSE,
  top_impact       = NULL,
  
  # === Variable Display ===
  var_name_by_description = TRUE,
  add_var_info            = FALSE,
  
  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 45,
    height = 20
  ),
  
  # === Styling ===
  plot_style_config = create_plot_style(
    positive_color = "#2E8B57",
    negative_color = "#CD5C5C",
    panel_rows = 1,
    panel_cols = NULL,
    show_axis_titles_on_all_facets = FALSE,
    y_axis_text_size = 25,
    bar_width = 0.6,
    all_font_size = 1.1
  )
)
```

<details>
  <summary class="toggle-summary">💡 Tip</summary>
  
<div class="tip-box">
  You may notice the following adjustments:
  
- `positive_color` and `negative_color`: Define colors for positive and negative values, automatically distinguishing impacts by shade.

- `show_axis_titles_on_all_facets`: Suppresses repeated labels across facets.

- `y_axis_text_size`: Increases y-axis label size to 25 from the default setting.

- `bar_width`: Decreases bar width for better visualization of large datasets.

- `all_font_size`: Increases all font sizes to 1.3 from the default setting.
</div>

</details>

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(plotB[[1]])
```
</div> <figcaption class="package-fig"> Detail Plot Visualization <br> <span class="package-fig-note">(top_impact = NULL, invert_axis = TRUE, panel_rows = 1)</span> </figcaption> </figure>
  
## Example 2  

Another powerful feature of `detail_plot` is the `top_impact` argument,  
which automatically filters and displays the most-affected values (both gains and losses)  
based on the `x_axis_from` grouping.

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.width=12, fig.height=10}
plotB2 <- detail_plot(
  # === Input Data ===
  data        = sector_data,
  filter_var  = list(Region = c("Oceania")),
  x_axis_from = "Commodity",
  split_by    = "Region",
  panel_var   = "Experiment",
  variable_col = "Variable",
  unit_col     = "Unit",
  desc_col     = "Description",
  
  # === Plot Behavior ===
  invert_axis      = TRUE,
  separate_figure  = FALSE,
  top_impact       = 10,
  
  # === Variable Display ===
  var_name_by_description = FALSE,
  add_var_info            = FALSE,
  
  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 25,
    height = 12
  ),
  
  # === Styling ===
  plot_style_config = create_plot_style(
    positive_color = "#2E8B57",
    negative_color = "#CD5C5C",
    panel_rows = 1,
    panel_cols = NULL,
    show_axis_titles_on_all_facets = FALSE,
    y_axis_text_size = 25,
    bar_width = 0.6,
    all_font_size = 1.1
  )
)

print(plotB2[[1]])
```
</div> <figcaption class="package-fig"> Detail Plot Visualization (filtered) <br> <span class="package-fig-note">(top_impact = 10, invert_axis = TRUE)</span> </figcaption> </figure>

---

## Example 3

Another example involves plotting multi-dimensional data like bilateral trade (`qxs`).  
To do this, specify:

- `split_by = c("Source", "Destination")`

This automatically generates separate plots for each unique pair, e.g., `"Oceania-SEAsia"`, `"Oceania-EastAsia"`, etc.  

Each plot shows detailed impacts at the sector (`COMM`) level for that country pair.

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
bilateral_data <- bilateral_data[["qxs"]]

plotB3 <- detail_plot(
  # === Input Data ===
  data        = bilateral_data,
  filter_var  = list(Source = c("Oceania"),
                     Destination = c("SEAsia", "EastAsia")),
  x_axis_from = "Commodity",
  split_by    = c("Source", "Destination"),    # The most important part!
  
  # === Plot Behavior ===
  invert_axis      = TRUE,
  separate_figure  = FALSE,
  top_impact       = NULL,
  
  # === Variable Display ===
  var_name_by_description = FALSE,
  add_var_info            = FALSE,
  
  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 45,
    height = 20
  ),
  
  # === Styling ===
  plot_style_config = create_plot_style(
    positive_color = "#2E8B57",
    negative_color = "#CD5C5C",
    panel_rows = 1,
    panel_cols = NULL,
    show_axis_titles_on_all_facets = FALSE,
    y_axis_text_size = 25,
    bar_width = 0.6,
    all_font_size = 1.1
  )
)
```

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(plotB3[[1]])
```
</div> <figcaption class="package-fig"> Detail Plot Visualization (filtered) <br> <span class="package-fig-note">(top_impact = 10, invert_axis = TRUE)</span> </figcaption> </figure>


# Stack Plot {#sec:stackplot}

This plot visualizes decomposition results, showing how different components contribute to a total value, making it particularly suitable for decomposition analysis.  

<strong style="font-size: 1.17em;">Key Features</strong>

- Displays the contribution of components to a total value.  

- Allows further decomposition by unstacking components instead of using a stacked bar.  

- Suitable for welfare decomposition, trade decomposition, and other GTAP structural breakdowns.  

<details>
  <summary class="toggle-summary">💡 Tip</summary>
  
  <div class="tip-box">
  Since decomposition data often contains multiple components, <strong>renaming and restructuring</strong> may be necessary before plotting. You can try the following code:
  </div>
  
```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
# Rename Value if needed
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

har.plot.data <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)
```
</details>

## Example 1

You can use the following command to plot the decomposition figure:

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
plotC <- stack_plot(
  # === Input Data ===
  data              = har.plot.data[["A"]],
  filter_var        = NULL,
  x_axis_from       = "Region",
  stack_value_from  = "COLUMN",
  split_by          = FALSE,
  panel_var         = "Experiment",
  variable_col      = "Variable",
  unit_col          = "Unit",
  desc_col          = "Description",

  # === Plot Behavior ===
  invert_axis     = FALSE,
  separate_figure = FALSE,
  show_total      = TRUE,
  unstack_plot    = FALSE,
  top_impact      = NULL,

  # === Variable Display ===
  var_name_by_description = TRUE,
  add_var_info            = FALSE,

  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 28,
    height = 15
  ),

  # === Styling ===
  plot_style_config = create_plot_style(
    color_tone                   = "gtap",
    panel_rows                   = 1,
    panel_cols                   = NULL,
    show_legend                  = TRUE,
    show_axis_titles_on_all_facets = FALSE
  )
)
```

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(plotC[[1]])
```
</div> <figcaption class="package-fig"> Stack Plot Visualization (stack) <br> <span class="package-fig-note">(split_by = FALSE, show_total = TRUE, unstack_plot = FALSE)</span> </figcaption> </figure>

## Example 2

To further analyze the decomposition at the component level, set `unstack_plot = TRUE` to display unstacked plots.

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.width=12, fig.height=10}
plotD <- stack_plot(
  # === Input Data ===
  data              = har.plot.data[["A"]],
  filter_var        = NULL,
  x_axis_from       = "Region",
  stack_value_from  = "COLUMN",
  split_by          = FALSE,

  # === Plot Behavior ===
  invert_axis     = TRUE,
  separate_figure = FALSE,
  show_total      = TRUE,
  unstack_plot    = TRUE,
  top_impact      = NULL,

  # === Variable Display ===
  var_name_by_description = TRUE,
  add_var_info            = FALSE,

  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 28,
    height = 15
  ),

  # === Styling ===
  plot_style_config = create_plot_style(
    color_tone                   = "gtap",
    panel_rows                   = 1,
    panel_cols                   = NULL,
    show_legend                  = TRUE,
    show_axis_titles_on_all_facets = FALSE
  )
)

print(plotD[[1]])
```
</div> <figcaption class="package-fig"> Stack Plot Visualization (unstacked) <br> <span class="package-fig-note">(split_by = FALSE, invert_axis = TRUE, unstack_plot = TRUE)</span> </figcaption> </figure>

## Example 3

Another practical example is plotting the decomposition of terms of trade (header E1), which is also a good example of multi-dimensional plotting. You may try the following command:

```{r, eval = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
plotE <- stack_plot(
  # === Input Data ===
  data              = har.plot.data[["E1"]],
  filter_var        = list(Region = c("Oceania", "SEAsia")),
  x_axis_from       = "Commodity",
  stack_value_from  = "PRICES",
  split_by          = "Region",
  panel_var         = "Experiment",
  variable_col      = "Variable",
  unit_col          = "Unit",
  desc_col          = "Description",

  # === Plot Behavior ===
  invert_axis     = TRUE,
  separate_figure = FALSE,
  show_total      = TRUE,
  unstack_plot    = FALSE,
  top_impact      = NULL,

  # === Export Settings ===
  output_path     = NULL,
  export_picture  = FALSE,
  export_as_pdf   = FALSE,
  export_config   = create_export_config(
    width  = 50,
    height = 30
  ),

  # === Styling ===
  plot_style_config = create_plot_style(
    title_format = create_title_format(
      type = "prefix",
      text = "Terms of Trade Decomposition",
      sep  = ": "
    ),
    color_tone                   = "winter", 
    panel_rows                   = 1,
    show_axis_titles_on_all_facets = FALSE,
    bar_width                    = 0.5,
    show_legend                  = TRUE
  )
)
```

<figure class="package-figure">
  <div class="package-figure-box">
```{r, echo=FALSE, fig.align='center', fig.width=12, fig.height=10}
print(plotE[[1]])
```
</div> <figcaption class="package-fig"> Stack Plot Visualization (Multi-Dimension) <br> <span class="package-fig-note">(split_by = "Region", plot_style_config -> adjustments)</span> </figcaption> </figure>


# Tip

## Plot Configurations

- Run the following chunks to view all available configurations.

```{r, eval=FALSE}
get_all_config()
```

- Rename the list outputs `my_export_config` and `my_style_config` to any names you prefer; you can create as many styles as needed.

- Enter the name of your custom style list into the plot function, while other settings in `(...)` remain unchanged, as shown in the sample below:

```{r, eval=FALSE}
comparison_plot <- (...,
                    export_config = my_export_config,
                    plot_style_config = my_style_config
                    )
```


## Sorting Output {#sec:sorting-output}

This function allows you to sort data before generating plots or tables.  
For example, you can display `EXP2` before `EXP1`, or `SEAsia` before `Oceania`.

It supports:

- Sorting by multiple columns simultaneously for precise output control  
- Applying consistent sorting across all data frames within a data list

See `?sort_plot_data` for full details.

Key arguments:  

- `cols` specifies the column(s) to sort based on a predefined order (e.g., in `sl4.plot.data`, `Experiment` and `Region` will follow the defined order).  

- `sort_by_value_desc` controls whether sorting is in descending order (`TRUE/FALSE/NULL`).   

To apply sorting, use the following command:  

```{r Sorting Data Function, eval = TRUE}
# Using the function
sort_sl4_plot <- sort_plot_data(
  sl4.plot.data,
  sort_columns = list(
    Experiment = c("EXP2", "EXP1"), # Column Name = Sorting Order
    Region = c("SEAsia", "Oceania")
    ),
  sort_by_value_desc = NULL
)
```

```{r, echo=FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.width=12, fig.height=10}
sort_sl4_plot = sort_sl4_plot[["REG"]]

plotE <- comparison_plot(
          # === Input Data ===
          data         = sort_sl4_plot,
          filter_var   = NULL,
          x_axis_from  = "Region",
          split_by     = "Variable",
          panel_var    = "Experiment",
          variable_col = "Variable",
          unit_col     = "Unit",
          desc_col     = "Description",
        
          # === Plot Behavior ===
          invert_axis     = FALSE,
          separate_figure = FALSE,
        
          # === Variable Display ===
          var_name_by_description = FALSE,
          add_var_info            = FALSE,
        
          # === Export Settings ===
          output_path     = NULL,
          export_picture  = FALSE,
          export_as_pdf   = FALSE,
          export_config   = create_export_config(
            width  = 20,
            height = 12
          ),
        
          # === Styling ===
          plot_style_config = create_plot_style(
            color_tone         = "purdue",
            add_unit_to_title  = TRUE,
            title_format       = create_title_format(
              type = "prefix",
              text = "Impact on"
            ),
            panel_rows = 2
          )
        )

print(plotE[[1]])
```


## Renaming Values {#sec:rename-value}

To automatically modify values in any column before plotting—such as renaming country names in the `REG` column (e.g., changing `"USA"` to `"United States"` or `"CHN"` to `"China"`)—use the following command:  

```{r Rename Value, eval = FALSE}
# Creating Mapping File
rename.region <- data.frame(
  ColumnName = "REG",
  OldName = c("USA", "CHN"),
  NewName = c("United States", "China"),
  stringsAsFactors = FALSE
)

sl4.plot.data_rename <- rename_value(sl4.plot.data, mapping.file = rename.region)
```

<br>
Another useful example is renaming the components of welfare decomposition before plotting.

```{r, eval = FALSE}
# Rename Value if needed
wefare.decomp.rename <- data.frame(
  ColumnName = "COLUMN",
  OldName = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
  NewName = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"),
  stringsAsFactors = FALSE
)

welfare_decomp <- rename_value(har.plot.data, mapping.file = wefare.decomp.rename)
```

# Sample Data

Sample data used in this vignette is obtained from the GTAPv7 model
and utilizes data from the GTAP 11 database. For
more details, refer to the [GTAP Database Archive](https://www.gtap.agecon.purdue.edu/databases/archives.asp).



```{=html}
<script>
window.addEventListener("load", function () {
  document.querySelectorAll("pre code").forEach(function (codeBlock) {
    const button = document.createElement("button");
    button.className = "copy-code-button";
    button.type = "button";
    button.innerText = "Copy";
    button.style.position = "absolute";
    button.style.top = "4px";
    button.style.right = "4px";

    button.addEventListener("click", () => {
      navigator.clipboard.writeText(codeBlock.innerText).then(() => {
        button.innerText = "Copied!";
        setTimeout(() => (button.innerText = "Copy"), 1500);
      });
    });

    const pre = codeBlock.parentNode;
    pre.style.position = "relative";
    pre.insertBefore(button, codeBlock);
  });
});
</script>

<script>
document.addEventListener("DOMContentLoaded", function () {
  let prevScrollPos = window.pageYOffset;
  const navbar = document.querySelector(".navbar");

  window.addEventListener("scroll", function () {
    const currentScrollPos = window.pageYOffset;
    if (prevScrollPos > currentScrollPos) {
      // Scrolling up
      navbar.style.top = "0";
    } else {
      // Scrolling down
      navbar.style.top = "-70px";
    }
    prevScrollPos = currentScrollPos;
  });
});
</script>
```
