---
title: "GTAPViz: Automating GTAP Data Processing with HARplus"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GTAPViz: Automating GTAP Data Processing with HARplus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)

required_pkgs <- c("tidyverse", "writexl", "dplyr", "devtools", 
                   "openxlsx", "readxl", "knitr", "rmarkdown", "data.table", 
                   "ggplot2", "tcltk", "gridExtra", "haven", "citation", "tinytex", "bookdown", "HARplus")

# Load packages quietly, but only if available
lapply(required_pkgs, function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    warning(sprintf("Package '%s' is not installed. Please install it before running this vignette.", pkg))
  }
})
# Directory
project.folder <- "D:/GitHub/GTAPViz/inst/extdata"
input.folder <- paste0(project.folder, "/in")
map.folder <- paste0(project.folder, "/map")
mapping.output <- system.file("extdata", "map", "OutputMapping.xlsx", package = "GTAPViz")
sl4map <- readxl::read_xlsx(mapping.output, sheet = "SL4File")
harmap <- readxl::read_xlsx(mapping.output, sheet = "HARFile")
setwd <- "D:/GitHub/GTAPViz/R"
devtools::load_all()

#devtools::install_github("Bodysbobb/GTAPViz")
```

# Introduction {#sec:introduction}

GTAPViz is built upon the **HARplus** package to facilitate the workflow of GTAP users by automating data extraction and processing from `.sl4` and `.har` result files. This vignette illustrates how to efficiently handle GTAP model output files using GTAPViz, ensuring seamless data manipulation for further analysis.

For data processing, the default variable definitions are based on **GTAP Model Version 7**, aligning with the standard structure used in the Global Trade Analysis Project (GTAP). Users can customize variable mappings, extract relevant indicators, and format the data for various analytical needs.

GTAPViz provides structured methodologies to process these datasets, integrating automated extraction, unit assignments, and data validation. By leveraging HARplus functionalities, this package simplifies handling multi-dimensional economic data from GTAP simulations.

For more details on HARplus, see:  
[\href{https://github.com/Bodysbobb/HARplus}{HARplus GitHub Repository}](https://github.com/Bodysbobb/HARplus).


# Project Directory {#sec:project-setting}

To use this R code, all solution files, including `.sl4` and `.har`, must be saved from RUNGTAP or CMD and placed in the same folder.

Adjust the `<project.folder>` directory to your desired path. If you maintain the same folder structure as in this code, you may not need to modify the input (`<input.folder>`), output (`<output.folder>`), or mapping (`<map.folder>`) directories.

By default, this R code assumes that `<project.folder>` contains the following three main folders:

-   **in** – Stores all input files (i.e., `.sl4` and `.har`). See the [Solution files](#sec:input-files) for details.
-   **out** – Stores all exported output files.
-   **map** – Stores the mapping `.xlsx` file (**critical for processing**).

<div style="font-family: monospace; background-color: #282C34; color: #ABB2BF; padding: 10px; border-radius: 5px; width: fit-content;">
📂 project.folder/ <br>
 ├── 📂 in/  <span style="color: #61AFEF;"># Stores input files (<code>.sl4</code>, <code>.har</code>)</span> <br>
 ├── 📂 out/ <span style="color: #98C379;"># Stores output files</span> <br>
 ├── 📂 map/ <span style="color: #E5C07B;"># Stores the mapping file (<code>OutputMapping.xlsx</code>)</span> <br>
</div>
   
: Example of Project Folder

# Mapping Files {#sec:mapping-files}

The **`<OutputMapping.xlsx>`** specifies the variables to extract from the solution files `<SL4File sheet>` and the HAR file `<HARFile sheet>`. All required variables must be listed in the `"Variable"` column.

For `"Description"` and `"Unit"`, you may:

- Leave the columns blank and put `"No"` in `<info.mode>` [setting](#sec:setting) to exclude these columns.  
  \footnote{If you leave them blank and put `"Yes"` instead of `"No"`, the result will return an empty column in your output.}

- Manually define **all variables** and put `"Yes"` in `<info.mode>` [setting](#sec:setting).

- Leave the columns blank and put `"GTAPv7"` in `<info.mode>` [setting](#sec:setting) to get the default definition and unit based on GTAPv7.  
  \footnote{The GTAP default can only define variables based on the GTAPv7 default set. If you add additional variables to the model, you must manually define the `"Description"` and `"Unit"` in the sheet if needed.}

- Define only some variables, leave others blank, and put `"Mix"` in `<info.mode>` [setting](#sec:setting) to get the definition based on your input for defined variables and the GTAP default for the others.


<div style="font-family: monospace; background-color: #282C34; color: #ABB2BF; padding: 10px; border-radius: 5px; width: fit-content;">
  <table style="border-collapse: collapse; width: 100%;">
    <tr>
      <th style="border: 1px solid #ABB2BF; padding: 8px; text-align: left;">Variable</th>
      <th style="border: 1px solid #ABB2BF; padding: 8px; text-align: left;">Description</th>
      <th style="border: 1px solid #ABB2BF; padding: 8px; text-align: left;">Unit</th>
    </tr>
    <tr>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">qgdp</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">Real GDP Index (%)</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">percent</td>
    </tr>
    <tr>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">EV</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">Welfare Equivalents</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">million USD</td>
    </tr>
    <tr>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">pebfactreal</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">Cost of endowment</td>
      <td style="border: 1px solid #ABB2BF; padding: 8px;">percent</td>
    </tr>
  </table>
</div>

: Example of SL4File and HARFile sheet


💡 **Tip:** The mapping file (`<OutputMapping.xlsx>`) can also be manually created in R as a dataframe, provided that it follows the exact structure, including column names. The dataframe must contain at least the `"Variable"` column, while `"Description"` and `"Unit"` columns are optional depending on the selected `<info.mode>`. 

```{r R Output Mapping, eval = FALSE}
mapping_df <- data.frame(
  Variable = c("qgdp", "EV", "pebfactreal"),
  Description = c("Real GDP Index (%)", "Welfare Equivalents", "Cost of endowment"),
  Unit = c("percent", "million USD", "percent"),
  stringsAsFactors = FALSE
)
```

# Input Files {#sec:input-files}

The `<case.name>` variable in [setting](#sec:setting) contains experiment names from which results will be extracted and merged into a single output for further analysis, such as graph creation. Multiple experiments can be included, but increasing their number will extend processing time.

## Input Files Format

All input files must be stored in `<input.folder>` (default: `"in"` inside `<project.folder>`). File names must follow `<case.name>` to ensure consistency between input each `.sl4` file must have a corresponding `-WEL.HAR` file

-   `EXP1.sl4`, `EXP1-WEL.har`
-   `EXP2.sl4`, `EXP2-WEL.har`

<div style="font-family: monospace; background-color: #282C34; color: #ABB2BF; padding: 10px; border-radius: 5px; width: fit-content;">
📂 in/ <br>
 ├── 📄 EXP1.sl4 <br>
 ├── 📄 EXP1-WEL.har <br>
 ├── 📄 EXP2.sl4 <br>
 ├── 📄 EXP2-WEL.har <br>
</div>
: Example of Input Files within the Input Folder

# Output Files {#sec:output-files}

This R code, by default, can generate four output formats: CSV, STATA, R, and Text files, defined by `<csv.output>`, `<stata.output>`, `<r.output>`, and `<txt.output>`, respectively. In [setting](#sec:setting), placing `"Yes"` or `"No"` for the required format. 

At the end, the code will automatically generate a `"README.xlsx"` file, reporting the location of exported key variables (i.e., sheet names). Since GTAP results contain different data dimensions, the output will be separated by dimension (1D, 2D, 3D).

For 1D and 2D data, results are automatically grouped into `"Region"` and `"Sector"`—distinguishing variables reported at the sector level from those reported by region—for easier data manipulation.

However, data larger than 3D will be reported separately by dimension, with output files named according to their original dimension names. An exception is `qxs`, which represents bilateral trade changes; if selected for export, it will be placed in the `"BilateralTrade"` sheet.


# Flexibility and Applicability {#sec:flexibility}

This R code is designed to extract any variable from `.SL4` and `.HAR` files, even those newly created by users that are not included in the original GTAP model. 

Furthermore, this code is adaptable to any version of the GTAP CGE model, provided that the `.SL4` and `.HAR` files exist and adhere to the required format.\footnote{With minor modifications, this R script can also be used to extract all types of three-dimensional data formatted in `.SL4` and `.HAR`.} 

In cases where `HARFile.HAR` files are unavailable, the script will still function, but the decomposition component will be skipped.

# Sample Data
Sample data used in this vignette is obtained from the GTAPv7 model and utilizes publicly available data from the GTAP 9 database. For more details, refer to the [GTAP Database Archive](https://www.gtap.agecon.purdue.edu/databases/archives.asp).

# Loading Package
Before proceeding, ensure that `GTAPViz` and `HARplus` are installed and loaded:
```{r package, eval = FALSE}
library(GTAPViz)
```

# Output Setup {#sec:setting}

This chunk defines the **experiment names**, **handling of description and unit columns**, and **output formats** for processing GTAP model results.

- **Experiment Names (`experiment`)**:  
  Specifies the names of experiments to process, matching the prefixes of `.sl4` and `.har` files.  
  Example:  
  - `EXP1.sl4` and `EXP1-WEL.har` → Processed as `EXP1`.  
  - `EXP2.sl4` and `EXP2-WEL.har` → Processed as `EXP2`.  

- **Description and Unit Handling (`info.mode`)**:  
  Determines how the `"Description"` and `"Unit"` columns are included in the output:  
  - `"Yes"` → Use manually defined descriptions and units from the mapping file.  
  - `"No"` → Exclude the `"Description"` and `"Unit"` columns.  
  - `"GTAPv7"` → Use default definitions and units based on **GTAP Model Version 7**.  
  - `"Mix"` → Use manually defined values where available, and GTAP defaults for the rest.  

- **Output Formats**:  
  Controls which file formats are generated (`"Yes"` = export, `"No"` = skip):  
  - **CSV (`csv.output`)** → `"No"`  
  - **STATA (`stata.output`)** → `"No"`  
  - **R (`r.output`)** → `"Yes"` (only `.rds` output enabled)  
  - **Text (`txt.output`)** → `"No"`  

This setup allows users to customize how data is processed and exported based on their needs.

```{r Input and Output Setup, eval = FALSE}
# Define experiment name / output name 
experiment <- c("EXP1", "EXP2")

# Adding Description / Unit (Yes/No/GTAPv7/Mix)
info.mode <- "GTAPv7"

# Output: (CSV, STATA, R) 
# "Yes" to opt for this output format, "No" to skip the output format
csv.output <- "No"    
stata.output <- "No"  
r.output <- "Yes"
txt.output <- "No"   
```

# Extracting Data 

```{r GTAP Automating, eval = FALSE}
process_gtap_data(
  sl4file = sl4map, 
  harfile = harmap,
  experiment = experiment,
  mapping_info = info.mode,
  project_dir = project.folder,
  export_data = FALSE
)
```

