---
title: "GTAPViz: Visualizing GTAP Model Results"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GTAPViz: Visualizing GTAP Model Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("GTAPViz", quietly = TRUE)
)

required_pkgs <- c("tidyverse", "writexl", "dplyr", "devtools", 
                   "openxlsx", "readxl", "knitr", "rmarkdown", "data.table", 
                   "ggplot2", "tcltk", "gridExtra", "haven", "citation", "tinytex", "bookdown", "HARplus")

# Load packages quietly, but only if available
lapply(required_pkgs, function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    warning(sprintf("Package '%s' is not installed. Please install it before running this vignette.", pkg))
  }
})

# Directory
project.folder <- "D:/One Drive/OneDrive - purdue.edu/GTAPViz Data/Plot"
input.folder <- paste0(project.folder, "/in")
output.folder <- paste0(project.folder, "/out")
map.folder <- paste0(project.folder, "/map")
mapping.output <- paste0(map.folder, "/OutputMapping.xlsx")
sl4plot <- readxl::read_xlsx(mapping.output, sheet = "SL4Plot")
harplot <- readxl::read_xlsx(mapping.output, sheet = "HARPlot")
setwd <- "D:/GitHub/GTAPViz/R"
devtools::load_all()

#devtools::install_github("Bodysbobb/GTAPViz")
```


# Introduction {#sec:introduction}

GTAPViz, built on the **HARplus** package, enhances the visualization of GTAP model results, making it easier for users to interpret and communicate economic simulation outputs. This vignette demonstrates how to create structured, publication-ready figures using GTAPViz’s built-in visualization tools.

The visualization framework relies on the **`SL4Plot`** and **`HARPlot`** sheets, which specify the variables and figure titles used for plotting. Users can define these directly within an external mapping file or generate them dynamically in the R environment.

By leveraging HARplus for data extraction and GTAPViz for visualization, this package provides a streamlined workflow to generate insightful graphs without requiring extensive manual coding.

For an overview of data processing functionalities, refer to the separate vignette:  
[\href{https://github.com/Bodysbobb/GTAPViz}{GTAPViz - Data Processing}](https://github.com/Bodysbobb/HARplus).

# Project Directory {#sec:project-setting}

To use this R code, all solution files, including `.sl4` and `.har`, must be saved from RUNGTAP or CMD and placed in the same folder.

Adjust the `<project.folder>` directory to your desired path. If you maintain the same folder structure as in this code, you may not need to modify the input (`<input.folder>`), output (`<output.folder>`), or mapping (`<map.folder>`) directories.

By default, this R code assumes that `<project.folder>` contains the following three main folders:

-   **in** – Stores all input files (i.e., `.sl4` and `.har`). See the [Solution files](#sec:input-files) for details.
-   **out** – Stores all exported output files.
-   **map** – Stores the mapping `.xlsx` file (**critical for processing**).

<div style="font-family: monospace; background-color: #282C34; color: #ABB2BF; padding: 10px; border-radius: 5px; width: fit-content;">
📂 project.folder/ <br>
 ├── 📂 in/  <span style="color: #61AFEF;"># Stores input files (<code>.sl4</code>, <code>.har</code>)</span> <br>
 ├── 📂 out/ <span style="color: #98C379;"># Stores output files</span> <br>
 ├── 📂 map/ <span style="color: #E5C07B;"># Stores the mapping file (<code>OutputMapping.xlsx</code>)</span> <br>
</div>
   
: Example of Project Folder

# Mapping Files {#sec:mapping-files}

The **`<OutputMapping.xlsx>`** specifies the variables to extract from the solution files `<SL4File sheet>` and the HAR file `<HARFile sheet>`. All required variables must be listed in the `"Variable"` column.

For `"Description"` and `"Unit"`, you may:

- Leave the columns blank and put `"No"` in `<info.mode>` [setting](#sec:setting) to exclude these columns.  
  \footnote{If you leave them blank and put `"Yes"` instead of `"No"`, the result will return an empty column in your output.}

- Manually define **all variables** and put `"Yes"` in `<info.mode>` [setting](#sec:setting).

- Leave the columns blank and put `"GTAPv7"` in `<info.mode>` [setting](#sec:setting) to get the default definition and unit based on GTAPv7.  
  \footnote{The GTAP default can only define variables based on the GTAPv7 default set. If you add additional variables to the model, you must manually define the `"Description"` and `"Unit"` in the sheet if needed.}

- Define only some variables, leave others blank, and put `"Mix"` in `<info.mode>` [setting](#sec:setting) to get the definition based on your input for defined variables and the GTAP default for the others.


| Variable     | Description         | Unit        |
|--------------|---------------------|-------------|
| qgdp         | Real GDP Index (%)  | percent     |
| EV           | Welfare Equivalents | million USD |
| pebfactreal  | Cost of endownment  | percent     |

: Example of SL4File and HARFile sheet


The `<SL4Plot sheet>` and `<HARPlot sheet>` are used for creating figure illustrations. Instructions for this can be found in another vignette called *GTAPViz_plot*. Essentially, the table contains two columns: `"Variable"`, which specifies the data used for figure creation, and `"PlotTitle"`, which defines the figure name. Users familiar with the R language can also create this table as a dataframe within the R environment.

| Variable      | PlotTitle             | Unit        |
|---------------|-----------------------|-------------|
| qgdp          | Real GDP Index        | percent     |
| EV            | Welfare Equivalents   | million USD |
| ppriv         | Consumer Price Index  | percent     |

: Example of SL4Plot and HARPlot sheet


💡 **Note:** The mapping file (`<OutputMapping.xlsx>`) can also be manually created in R as a dataframe, provided that it follows the exact structure, including column names. The dataframe must contain at least the `"Variable"` column, while `"Description"` and `"Unit"` columns are optional depending on the selected `<info.mode>`.  


For example, the following R code creates an equivalent mapping dataframe:

```{r, eval=FALSE}
mapping_df <- data.frame(
  Variable = c("qgdp", "EV", "ppriv"),
  PlotTitle = c("Real GDP Index", "Welfare Equivalents", "Cost of endowment"),
  Unit = c("Percent", "million USD", "percent"),
  stringsAsFactors = FALSE
)
```

# Input Files {#sec:input-files}

The `<case.name>` variable in [setting](#sec:setting) contains experiment names from which results will be extracted and merged into a single output for further analysis, such as graph creation. Multiple experiments can be included, but increasing their number will extend processing time.

## Input Files Format

All input files must be stored in `<input.folder>` (default: `"in"` inside `<project.folder>`). File names must follow `<case.name>` to ensure consistency between input each `.sl4` file must have a corresponding `-WEL.HAR` file

-   `EXP1.sl4`, `EXP1-WEL.har`
-   `EXP2.sl4`, `EXP2-WEL.har`

<div style="font-family: monospace; background-color: #282C34; color: #ABB2BF; padding: 10px; border-radius: 5px; width: fit-content;">
📂 in/ <br>
 ├── 📄 EXP1.sl4 <br>
 ├── 📄 EXP1-WEL.har <br>
 ├── 📄 EXP2.sl4 <br>
 ├── 📄 EXP2-WEL.har <br>
</div>
: Example of Input Files within the Input Folder

# Sample Data
Sample data used in this vignette is obtained from the GTAPv7 model and utilizes publicly available data from the GTAP 9 database. For more details, refer to the [GTAP Database Archive](https://www.gtap.agecon.purdue.edu/databases/archives.asp).

# Loading Package
Before proceeding, ensure that `GTAPViz` and `HARplus` are installed and loaded:
```{r package}
library(GTAPViz)
```

# Input Setup {#sec:setting}

This chunk defines the **experiment names**, **handling of description and unit columns**, and **output formats** for processing GTAP model results.

- **Experiment Names (`experiment`)**:  
  Specifies the names of experiments to process, matching the prefixes of `.sl4` and `.har` files.  
  Example:  
  - `EXP1.sl4` and `EXP1-WEL.har` → Processed as `EXP1`.  
  - `EXP2.sl4` and `EXP2-WEL.har` → Processed as `EXP2`.  

- **Description and Unit Handling (`info.mode`)**:  
  Determines how the `"Description"` and `"Unit"` columns are included in the output:  
  - `"Yes"` → Use manually defined descriptions and units from the mapping file.  
  - `"No"` → Exclude the `"Description"` and `"Unit"` columns.  
  - `"GTAPv7"` → Use default definitions and units based on **GTAP Model Version 7**.  
  - `"Mix"` → Use manually defined values where available, and GTAP defaults for the rest.  
  
```{r Input Setup}
# Define experiment name / output name 
case.name <- c("US_All", "US_All_RetalTar", "US_All_ReduceTar50", "US_All_RegReduceTar50",
                    "US_All10", "US_All10_RetalTar", "US_All10_ReduceTar50", "US_All10_RegReduceTar50")

# Adding Description / Unit (Yes/No/GTAPv7/Mix)
info.mode <- "GTAPv7"
```


# Preparing Input Data
```{r Preparing Data for Plot}
# Define variables to extract
# Region to be plotted
selected_regions <- c("USA", "CHN", "CAN", "ASEAN", "ROW")

# Sector to be plotted (NULL to select all)
selected_sector <- NULL

# Extract data with region and experiment filters
plot.dta <- plot_gtap_data(
  sl4file  = sl4plot,
  harfile = harplot, 
  experiment  = case.name,
  mapping_info = info.mode, 
  region_select  = selected_regions,
  sector_select = selected_sector,
  project_dir = project.folder
)

# Convert Value if Needed
sl4.plot.data <- convert_units(
  sl4.plot.data,
  change_unit_from = c("million USD"),
  change_unit_to = c("billion USD"),
  adjustment = c("/1000")
)

har.plot.data <- convert_units(
  har.plot.data,
  change_unit_from = c("million USD"),
  change_unit_to = c("billion USD"),
  adjustment = c("/1000")
)

# Convert the column name if needed
rename_col <- data.frame(
  old = c("REG", "COMM", "ACTS"),
  new = c("Region", "Commodity", "Activity")
  )
sl4.plot.data <- HARplus::rename_dims(sl4.plot.data, rename_col)
har.plot.data <- HARplus::rename_dims(har.plot.data, rename_col)

# Define the sublists to keep in the first list
reg_dim <- c("qgdp", "ppriv", "EV", "tot", "u")
reg_sl4 <- sl4.plot.data[reg_dim]
reg_sl4_map <- sl4plot[sl4plot$Variable %in% reg_dim, ]
reg_oth_sl4 <- sl4.plot.data[setdiff(names(sl4.plot.data), reg_dim)]
reg_oth_map <- sl4plot[!sl4plot$Variable %in% reg_dim, ]

```

# Comparison Plot
```{r Comparison Plot}
lapply(reg_sl4, function(df) {
  comparison_plot(df,
                  x_axis_from = "Region",
                  plot_var = reg_sl4_map,
                  title_prefix = "Impact on",
                  output_dir = output.folder,
                  compare_by_x_axis = FALSE,
                  separate_figure = FALSE,
                  color_tone = "grey",
                  width = 20,
                  height = 12,
                  legend_position = "bottom")})

report_table(reg_sl4, vars = reg_sl4_map$Variable, 
             col_names = reg_sl4_map$PlotTitle,  
             x_axis_from = "Region",
             output_dir = output.folder)
```

# GTAP MAcro Variable Plot (i.e., Scalar Plot without Region)
```{r GTAP Macro Plot}
Macros <- do.call(gtap_macros_data, c(as.list(paste0(input.folder, "/", case.name, ".sl4")),
                  list(experiment_names = case.name)))

# Selecting Variables
selected_marcro <- c("pgdpwld", "qgdpwld", "vgdpwld")
macros_new_name <- c("pgdpwld", "qgdpwld", "vgdpwld")
macro.map <- data.frame(
  Variable = selected_marcro,
  PlotTitle = macros_new_name,
  stringsAsFactors = FALSE
)

# Plotting
macro_plot(Macros,
           plot_var = macro.map,
           compare_by_x_axis = FALSE,
           color_tone = "grey",
           output_dir = output.folder,
           panel_rows = 2,
           width = 45,
           separate_figure = FALSE,
           legend_position = "bottom")

# Reporting Table
scalar_table(Macros, vars = selected_marcro, 
             output_dir = output.folder)
```

# Detail Plot
```{r Detail Plot}
lapply(reg_oth_sl4, function(df) {
  detail_plot(df,
              y_axis_from = c("Commodity", "Activity"),
              figure_separate_by = "Region",
              plot_var = reg_oth_map,
              output_dir = output.folder,
              top_impact = NULL,
              panel_rows = 1,
              panel_cols = 8,
              width = 45)})

```

# Decomposition Plot
```{r Decomposition Plot}
# Rename Value if needed
welfare <- har.plot.data[["A"]]
welfare[["Variable"]] <- "Welfare"
welfare <- rename_value(har.plot.data[["A"]], 
                        column_name = "COLUMN",
                        original_values = c("alloc_A1", "ENDWB1", "tech_C1", "pop_D1", "pref_G1", "tot_E1", "IS_F1"),
                        new_values = c("Alloc Eff.", "Endwb", "Tech Chg.", "Pop", "Perf", "ToT", "I-S"))

# Define Decomposition
stack_plot(data = welfare,
          x_axis_from = "Region",
          stack_value_from = "COLUMN",
          title_prefix = "Decomposition",
          output_dir = output.folder,
          separate_figure = FALSE,
          color_tone = "green",
          width = 30,
          height = 15,
          legend_position = "bottom",
          unstack_plot = FALSE)

# Table of Decomposition
decomp_table(har.plot.data, 
             header = c("A", "E1"),
             wide_cols = list(A = "COLUMN", E1 = "PRICES"),
             total_column = TRUE,
             output_dir = output.folder,
             export_table = TRUE)
```

